Windowed least-squares model reduction for dynamical systems

### 2019-10-16 section4/spatial_numerical.tex
through 4.2

Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time instances.  

@@

denotes the time step, $\lmsWidthArg{n}{i}$ with $\lmsWidthArg{1}{1} = 1$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window  and

@@

To describe the Gauss--Newton method, we first define a ``vectorization" function 
\begin{align*}
 \unroll &\vcentcolon (\stateyDiscreteArg{1},\ldots,\stateyDiscreteArg{m} ) \mapsto \begin{bmatrix} [\stateyDiscreteArg{1}]^T & \ldots & [\stateyDiscreteArg{m}]^T \end{bmatrix}^T  \\
&\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm}.
\end{align*}
The vectorized generalized coordinates over the $n$th time window are then
defined as \KTC{notationally, why are we doing a double bar instead of a bar?}
\EP{The notation for space--time in 2.4 uses bars. I don't want to confuse these generalized coordinates with those, they conceptually two different things} 
\begin{equation*}
\genstatecollocMatSlabArg{n} \defeq 
\unroll (\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}} ).
\end{equation*}
We now define the weighted space--time residual over the entire window as
\KTC{To make this valid for seeping back into the previous time window, the
second argument could be omitted here (okay because there is already a
superscript on the residual) and arbitrary previous solutions could be
included in the first residual here.}
\begin{equation*}
\residLMSSlabArg{n} : (\genstatecollocMatySlabArg{}) \mapsto \begin{bmatrix}
 \sqrt{\frac{\quadWeightsLMSScalarArg{n}{1}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n}; 
%\residLMSArg{n,0}( \basisspaceArg{n} \genstateyDiscreteArgnt{0} + \stateInterceptArg{n}),
\projectorSpatialArg{n}( \approxstateDiscreteArg{n,0}),
\ldots , \projectorSpatialArg{n}( \approxstateDiscreteArg{n,1 - \lmsWidthArg{n}{1} } )  ) \\
\vdots \\
 \sqrt{\frac{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}; \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}-1} + \stateInterceptArg{n} ,  \ldots,\projectorSpatialArg{n}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - \lmsWidthArg{n}{\nstepsArg{n}} } + \stateInterceptArg{n} )) \\
\end{bmatrix},
\end{equation*}
where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
It is worth noting that, by design, 
\begin{equation*}
\objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
=
\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}) \bigg].
\end{equation*} 
Using these definitions, Algorithm~\ref{alg:colloc_gn} presents the standard Gauss--Newton method. The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the Jacobian of the residual over the time window and form the normal equations, and (3) solve the normal equations and update the state.

>>>>

Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $\lmsWidthArg{n}{i}$ time instances, where $\lmsWidthArg{n}{i}$ (with $\lmsWidthArg{1}{1} = 1$) denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window. 

@@

denotes the time step, and

@@

To describe the Gauss--Newton method, we first define a ``vectorization" function 
\begin{align*}
 \unroll &\vcentcolon (\stateyDiscreteArg{1},\ldots,\stateyDiscreteArg{m} ) \mapsto \begin{bmatrix} [\stateyDiscreteArg{1}]^T & \ldots & [\stateyDiscreteArg{m}]^T \end{bmatrix}^T  \\
&\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm}.
\end{align*}
The vectorized generalized coordinates over the $n$th time window are then
defined as \KTC{notationally, why are we doing a double bar instead of a bar?}
\EP{The notation for space--time in 2.4 uses bars. I don't want to confuse these generalized coordinates with those, they conceptually two different things} 
\begin{equation*}
\genstatecollocMatSlabArg{n} \defeq 
\unroll (\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}} ).
\end{equation*}
We now define the weighted space--time residual over the entire window as
\KTC{To make this valid for seeping back into the previous time window, the
second argument could be omitted here (okay because there is already a
superscript on the residual) and arbitrary previous solutions could be
included in the first residual here.}
\begin{equation*}
\residLMSSlabArg{n} : \genstatecollocMatySlabArg{} \mapsto \begin{bmatrix}
 \sqrt{\frac{\quadWeightsLMSScalarArg{n}{1}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n}; 
%\residLMSArg{n,0}( \basisspaceArg{n} \genstateyDiscreteArgnt{0} + \stateInterceptArg{n}),
\projectorSpatialArg{n}( \approxstateDiscreteArg{n,0}),
\ldots , \projectorSpatialArg{n}( \approxstateDiscreteArg{n,1 - \lmsWidthArg{n}{1} } )  ) \\
\vdots \\
 \sqrt{\frac{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}; \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}-1} + \stateInterceptArg{n} ,  \ldots,\projectorSpatialArg{n}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - \lmsWidthArg{n}{\nstepsArg{n}} } + \stateInterceptArg{n} )) \\
\end{bmatrix},
\end{equation*}
where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
It is worth noting that, by design, 
\begin{equation*}
\objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
=
\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}) \bigg].
\end{equation*} 
Using these definitions, Algorithm~\ref{alg:colloc_gn} presents the standard Gauss--Newton method. The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the Jacobian of the residual over the time window and form the normal equations, and (3) solve the normal equations and update the state.
